The Emergent Simplicity of Galaxy Size

### 2017-09-28 paper/galsize_paper.tex
Updated abstract

\title[Empirical Modeling Galaxy Size]
{Empirical Modeling Galaxy Size Throughout the Cosmic Web}

@@

\begin{abstract}
We derive empirical modeling constraints on the connection between dark matter halos and the half-light radius $\rhalf$ of galaxies. Using forward-modeling techniques based on {\tt Halotools}, we confirm previous results in Kravtsov (2013) that $\rhalf$ is well-described by a linear scaling relation with halo virial radius. Novel to this work, we use new SDSS measurements of the $\rhalf-$dependence of galaxy clustering to test this modeling assumption. With no changes to the parameters, the model accurately predicts the observed two-point clustering on small- and large-scales over a wide range of stellar mass. This quantitative success is remarkable since the Kravtsov (2013) parameters were fit to the observed $\langle\rhalf\vert\mstar\rangle,$ and the $\rhalf-$dependence of SDSS galaxy clustering has heretofore never been measured. Moreover, this success is non-trivial, as we demonstrate that galaxy clustering is highly sensitive to the assembly history-dependent physics that shapes the relative size of centrals and satellites. Our results can be treated as a boundary condition for more complex and fine-grained models of galaxy size, and provide a simple means for cosmological surveys to generate synthetic galaxy populations with realistic sizes across the cosmic web.
\end{abstract}

>>>>

\title[The Emergent Simplicity of Galaxy Size]
{The Emergent Simplicity of Galaxy Size}

@@

\begin{abstract}
We derive empirical modeling constraints on the connection between dark matter halos and the half-light radius $\rhalf$ of galaxies. Using forward-modeling techniques based on {\tt Halotools}, we confirm previous results in Kravtsov (2013) that $\rhalf$ is well-described by a linear scaling relation with halo virial radius. Novel to this work, we study galaxy size using new SDSS measurements of the $\rhalf-$dependence of galaxy clustering. Smaller galaxies cluster more strongly relative to larger galaxies of the same stellar mass, a new result. We show that the $\rhalf-$dependence of galaxy clustering is largely driven by the relative size of central and satellite galaxies: satellites are smaller than centrals of the same halo mass. We identify the ingredients that contribute most significantly to $\rhalf-$dependent clustering by building empirical models of galaxy size with various effects included and excluded. Stellar mass stripping of satellites has only a mild impact on $\rhalf-$dependent clustering and is unlikely to be the origin of the characteristic shape of the signal. On the other hand, correlations between galaxy size and host halo assembly history leave a dramatic imprint on the clustering signal that is strikingly similar in magnitude, $\mstar-$dependence and scale-dependence. In the class of models that give the most faithful recovery of the observed clustering signal, galaxy $\rhalf$ is set by the size of the  virial radius at the time the subhalo reaches its peak halo mass, with log-normal scatter in $\rhalf$ that is strongly correlated with the size of the halo scale radius at that time. Our results can be treated as a boundary condition for more complex and fine-grained models of galaxy size, and provide a simple means for cosmological surveys to generate synthetic galaxy populations with realistic sizes across the cosmic web.
\end{abstract}
